\begin{abstract}

We introduce \textbf{enmeshed networks}, a new architectural primitive in which a lightweight parametric module shares the forward pass of a frozen host model, reading and writing intermediate representations at each layer while maintaining separate parameters and full removability. Unlike adapters (which require gradient descent per adaptation), side-tuning (which fuses only at the output layer), or activation steering (which applies fixed, non-adaptive directions), enmeshed networks create a \emph{parallel context channel}: a private subcontext processed through the same host model and compiled into modulation tensors that shape the primary context's computation without competing for its attention budget.

We instantiate this primitive as \textbf{HEXIS} (Hidden Enmeshed eXperiential Identity States, a name retained for the system; the technical claim is operational rather than cognitive). HEXIS implements compiled stance memory for frozen transformers via low-rank Q/V modulation. A compiled variant encodes a structured memory schema---the Mind Tree---into persistent modulation tensors at zero token cost, producing persistent behavioural stance biases that are dilution-immune by construction. On Qwen3.5-4B-Instruct, HEXIS achieves 100\% stance maintenance through 4K tokens of adversarial dilution and 83\% $\geq$4-conviction hold rate across 5 escalating pressure levels of \emph{illegitimate} adversarial pressure (fabricated counter-evidence, authority appeals, emotional pressure; vs.\ 0\% for context-level conditions, $n$=72 per cell), with 65\% shorter generation. We measure resistance to adversarial pressure rather than general updating behaviour; the protocol does not test whether the model updates appropriately on legitimate corrections. A three-layer architecture combining compiled enmeshment with M-curated belief slots and recursive expansion matches full in-context beliefs at 82\% token savings. The mechanism generalises beyond disposition to agentic tool orchestration: the same hidden-state bridge that compiles beliefs into attention modulation also serves as a retrieval query over a persistent knowledge graph. A contrastive retrieval phi ($\phi_R$) compiled in $\sim$20 seconds achieves 100\% R@1 on 108-node graphs, enabling a teacher loop where failure observations are written to the graph and retrieved on subsequent attempts. On the multi-domain $\tau^3$-bench (4 domains, 24-task balanced primary panel, 3 arms $\times$ 5 trials, 3-judge audited), HEXIS with teacher + always-on $d^*$ verify lifts pass rate by $+10$pp (61.7\% $\to$ 71.7\% excluding two airline tasks with documented LLM-judge variance, paired McNemar $p=0.043$ unadjusted, $p_\text{adj}=0.086$ Bonferroni); the strongest per-domain lift is on banking\_knowledge ($+20$pp). A LoRA fine-tune control on a 1{,}000-sample subset of AReaL-tau2 SFT data (covering airline, retail, telecom; banking\_knowledge \emph{not} in corpus) wins on its trained domains ($85$\% balanced vs HEXIS-C5 $80$\%) but fails entirely on the held-out domain (0\% --- the model emits malformed tool-call JSON), illustrating that HEXIS's contribution is deploy-time generalisation to unseen domains rather than peak performance on a known distribution. Cross-host scale validation on Qwen3.6-27B reproduces the stance-resistance lift at substantial magnitude ($+24$pp hold rate); cross-family transfer to Mistral-Ministral-8B is bounded by host RLHF refusal training ($+3$pp hold, $-15$pp cap rate). We identify two M mechanisms---attention modulation (mechanism a, for disposition) and knowledge retrieval/injection (mechanism b, for agentic tasks)---unified through the host model's hidden-state space. We characterise the bottleneck boundary precisely: compiled enmeshment steers parametric knowledge but cannot inject novel content. Code and checkpoints: \url{https://github.com/dgonier/hexis-public}.

\end{abstract}
