\section{Conclusion}
\label{sec:conclusion}

We introduced enmeshed networks---a new architectural primitive in which a lightweight module shares the forward pass of a frozen host, reading and writing intermediate representations at each layer---and instantiated this primitive as HEXIS, implementing compiled stance memory for frozen transformers via low-rank Q/V modulation.

HEXIS opens a parallel context channel: a structured memory schema (the Mind Tree) is compiled into persistent modulation tensors that shape the host's attention geometry without occupying context positions. The compiled signal is dilution-immune by construction (100\% stance at 4K tokens of filler), provides resistance to illegitimate adversarial pressure (83\% $\geq$4-conviction hold rate over 5 escalating pressure levels on the 4B instruct model vs.\ 0\% for non-compiled conditions; orthogonal to the question of whether the model updates correctly on legitimate corrections, which we do not test), and carries sufficient content through a rank-16 bottleneck to flip the host model's default stance on contested topics. A three-layer architecture (compiled modulation + curated slot + recursive expansion) matches full in-context beliefs at 82\% token savings.

We characterized the bottleneck boundary precisely: compiled enmeshment steers parametric knowledge but cannot inject novel content. This honest boundary motivates architectural composition rather than a single mechanism. The enmeshed modulation succeeds for open generation and fails for constrained action selection, establishing complementary roles for implicit and explicit memory channels.

The mechanism's value is sharpest in domains where knowledge is better encoded as an attention reweighting than as text: user-preference tracking that persists without per-turn re-prompting, research stance that resists adversarial consensus pressure, agentic strategy that does not dilute as episodes accumulate, and codebase awareness that scales at fixed rather than linear cost. In each case, the critical property is not token savings but the qualitative shift from inspectable content (which dilutes and folds) to non-inspectable modulation (which is structurally immune to both).

We view this work as opening a research program rather than presenting a final architecture. The enmeshed network primitive and the Mind Tree schema are general; the specific instantiation (Level 1, rank 16, Q+V, stride-3, Qwen3.5-4B-Base) is one point in a large design space. We hope the definition, taxonomy, and initial validation encourage exploration of the full space.
