\section{HEXIS: Architecture and Training}
\label{sec:hexis}

We instantiate a Level 1 enmeshed network (\S\ref{sec:enmeshed}) over Qwen3.5-4B-Base, a hybrid-attention model with 32 layers (24 DeltaNet linear attention, 8 full self-attention; $d=2560$, GQA with 16 query and 4 key-value heads). We term this instantiation HEXIS: the compiled modulation tensors are the \emph{states}---Hidden from the host's introspection, Enmeshed in its forward pass, derived from eXperience, and shaping its Identity.

\subsection{Modulation Mechanism}
\label{sec:modulation}

At each patched layer $\ell \in \mathcal{L}$, HEXIS applies two low-rank perturbations to the host's projections. For Q-modulation, the pre-projection hidden state $x_\ell$ is perturbed before the query projection:
\begin{equation}
x'_\ell = x_\ell + s_M \cdot (x_\ell \, M_A^\ell) (M_B^\ell)^\top, \qquad
V'_\ell = V_\ell + s_E \cdot (x_\ell \, E_A^\ell) (E_B^\ell)^\top
\label{eq:qv_mod}
\end{equation}
where $M_A^\ell, M_B^\ell, E_A^\ell, E_B^\ell$ are the rank-$r$ ($r=16$) modulation matrices and $s_M, s_E$ are learned scales. Q-modulation controls \emph{what the host attends to}; V-modulation controls \emph{what information is extracted} from attended positions. Together they implement a compiled attentional policy. We patch stride-3 across the network (11 of 32 layers, including 4 of 8 full-attention layers). Each modulated layer adds $\sim$82K FLOPs/token, $\sim$200$\times$ cheaper than adding equivalent belief tokens to context. A diagram of the two perturbations is in Appendix Figure~\ref{fig:modulation}.

\textbf{Query-agnostic compilation, content-specific effect.} The modulation tensors are compiled once from the Mind Tree and fixed; they encode the agent's disposition rather than a response to any query. Yet the \emph{effect} is query-specific because $x M_A M_B^\top$ depends on the current hidden state. M acts as a \emph{content-addressable filter}: each belief occupies different directions in the rank-$r$ subspace, and the input determines which directions are expressed.

\subsection{The Write Function $\phi$}
\label{sec:write_function}

$\phi$ compresses hidden states into modulation matrices. Given experience tokens processed through the host's frozen layers, $\phi$ reads the hidden states at each patched layer, pools them, and projects through a bottleneck: $z_\ell = \text{SiLU}(W_\text{down}^\ell \bar{h}_\ell)$, $(M_A^\ell, M_B^\ell) = \text{reshape}(W_\text{up}^\ell z_\ell)$, where $\bar h_\ell$ is the pooled hidden state at layer $\ell$. In compiled mode, pooling is weighted by conviction level. $\phi$ is trained jointly with the modulation objective and frozen at inference. Different inputs to $\phi$ produce different modulation tensors --- this is what makes enmeshed modulation experience-adaptive. A new belief set requires only one forward pass through $\phi$ ($\sim$20 seconds), so the agent can recompile M in seconds and immediately benefit from updated disposition. Full compilation pipeline (PhiNodeWriter, BeliefTreeMemory, MStateReadHead) in Appendix~\ref{sec:system_figures} (Figure~\ref{fig:system_overview}).

\subsection{Training and the Directional Channel $d^*$}
\label{sec:training}

Training follows a multi-phase curriculum (full details in Appendix~\ref{sec:training_curriculum}). Phase A teaches $\phi$ to encode beliefs and M to produce useful Q-modulation; Phase B trains the VModulationProjector to carry belief \emph{content} through the rank bottleneck so beliefs need not appear in the prompt; Phase D fixes verbatim repetition under sycophancy pressure using 44 gold challenge-response pairs. Total: $\sim$500 epochs, $\sim$10 hours on a single GPU.

HEXIS composes with representation engineering via the directional channel $d^*$, a fixed activation vector extracted from contrastive activation means: $d^* = N^{-1} \sum_i (\bar h_i^\text{pro} - \bar h_i^\text{con})$ over 62 topics. $d^*$ is injected as a post-attention residual at decode time. The critical property is that $d^*$ and M operate on orthogonal axes: the directional delta along $d^*$ between M-conditioned and unconditioned generation is $-0.001$ nats. Direction comes from $d^*$, disposition from M, content from the Mind Tree --- the three channels are architecturally separated.
